---
title: "Machine Learning course Project-HAR"
author: "Behzad Najafi"
date: "27 December 2015"
output: html_document
---
Introduction
----------------------------------------------------------------------------------------
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

Data
----------------------------------------------------------------------------------------
Both the testing and the training data can be downloaded from the following links
training data: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

test data: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

After downloading the files, the training and test data can be obtained as follows:
```{r}
raw_data<-read.csv("pml-training.csv")
finalTestData<-read.csv("pml-testing.csv")
```
Then we investigate the number of samples and columns of the available data as:
```{r}
dim(raw_data)
```
So the training data includes19622 samples with 160 parameters. The parameter to be estimated is the "classe" which, as demonstrated below can belong to one of the follwoing five categories:
```{r}
table(raw_data$classe)
```
Diving the raw data
----------------------------------------------------------------------------------------
Before going further, we should divide the available raw data into training and testing categories, as also demonstrated in the class, we can do so using the following commands
```{r}
#library(caret)
inTrain = createDataPartition(y=raw_data$classe, p=0.7, list=FALSE)
training = raw_data[inTrain,]
testing = raw_data[-inTrain,]
```
When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
