---
title: "Machine Learning course Project-HAR"
author: "Behzad Najafi"
date: "27 December 2015"
output: html_document
---
Introduction
----------------------------------------------------------------------------------------
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

Data
----------------------------------------------------------------------------------------
Both the testing and the training data can be downloaded from the following links
training data: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

test data: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
After downloading the files, the training and testing datasets can be obtained:
```{r}
raw_data<-read.table("pml-training.csv",header = TRUE, sep = ",", na.strings = c("NA", "#DIV/0!"))
finalTestData<-read.table("pml-testing.csv",header = TRUE, sep = ",", na.strings = c("NA", "#DIV/0!"))
```
Next,we investigate the number of samples and columns of the available data as:
```{r}
dim(raw_data)
```
So the training data includes 19622 samples with 160 parameters. The parameter to be estimated is the "classe" which, as demonstrated below, can belong to one of the follwoing five categories:
```{r}
table(raw_data$classe)
```
Cleaning the Data
----------------------------------------------------------------------------------------
Several  parameters include many NA values which should be removed, furthermore, the first seven parameters do not include valuable information for creating the model. Hence, the columns with mroe than 90% of NA values together with the first seven columns are eliminated.
```{r}
is_na_result =is.na(raw_data)
removedColumns = which(colSums(is_na_result) >0.9*dim(raw_data)[1])
clean_data1 = raw_data[, -removedColumns]
clean_data = clean_data1[, -c(1:7)]
dim(clean_data)
```
Diving the raw data
----------------------------------------------------------------------------------------
Now we have a clean dataset with 19622 samples and 53 columns.Before going further, we should divide the available raw data into training and testing (cross validation) categories:
```{r}
library(caret)
inTrain = createDataPartition(y=raw_data$classe, p=0.7, list=FALSE)
training = clean_data[inTrain,]
testing = clean_data[-inTrain,]
dim(training);dim(testing);
```
Building the Model
----------------------------------------------------------------------------------------
In this step we build the model using random forests: 
```{r}
library(randomForest)
fittedModel = train(classe~., method="rf", data=training)
```
Evlauting the Model
----------------------------------------------------------------------------------------
Using the trained model, we can now predict the parameter "classe" for the testing(cross validation)data set:
```{r}
predictions<-predict(fittedModel, testing)
mean(predictions==testing$classe)
```
The accuracy of the results can be determined using the confusionMatrix, comapring the predicted values and the real values of the objective parameter in the testing dataset:
```{r}
confusionMatrix(predictions, testing$classe)
```
An acceptable specificty and sensitivty can be observed for all of the classes. Finally, In order to predict the result of the final testing dataset:
```{r}
predictions_Finaltesting<-predict(fittedModel, finalTestData)
result<-as.character(predictions_Finaltesting)
```
Following vector demonstrates the obtained results:
[B A B A A E D B A A B C B A E E A B B B]